---
title: "Task1"
author: "Lucia Bušová"
date: "27/10/2025"
format:
  html:
    embed-resources: true
editor: source
---

## Question 1 - atomic types


Question description

```{r}
# code for solving task
#numeric
population <- 850000
size <- 320 
temperature <- 25
rating <- 5
# ineger
number_of_moons <- 2
buildings <- 4572
inhabitants <- 850000
# logical
has_water <- TRUE
has_air <- TRUE
# character
name <- "Lumeria"
region <- "emerald_valley_region"
category <- "Nature_oasis_city"


```

## Question 2 Vectors

Question description

```{r}
# code for solving task
#numeric vector
num_vector <- c(population,size,temperature,rating,number_of_moons, buildings,inhabitants)

#sequence vector
s_vector <- seq(from = 20, to = 100, by =10)

#logical vector
l_vector <- sample(c(TRUE,FALSE),size=10, replace = TRUE)
```

## Question 3 Matrix

Question description

```{r}
# code for solving task
x <- sample(-50:50, 9, replace = TRUE)
m <- matrix(x, nrow = 3, ncol = 3)
rownames(m) <- c("A1","A2","A3")
colnames(m) <- c("B1","B2","B3")

row_means <- rowMeans(m)
col_means <- colMeans(m)

```

## Question 4 List

Question description

```{r}
# code for solving task
T_list <- list(population=population,
               size=size,
               temperature=temperature,
               rating=rating,
               number_of_moons=number_of_moons,
               buildings=buildings,
               inhabitants=inhabitants,
               
               s_vector=s_vector <- seq(from = 15, to = 100, by =5),
               l_vector=l_vector <- sample(c(TRUE,FALSE),size=10, replace = TRUE),
               m=m,
               T_descript = 'Lumeria - oáza pokoja, v prírodnom prostredí')

```


## Question 5 Factor

Question description

```{r}
# code for solving task
development_level <- factor(c('very low','low', 'medium','high','very high'), levels = c('very low','low', 'medium','high','very high'))

table(development_level)
```

## Question 6 Data frame

Question description

```{r}
# code for solving task
cities_data <- data.frame(
  cities = c('Lumeria', 'Bratislava', 'Nitra', 'Trnava', 'Trenčín', 'Košice'),
  population = c(850000, 478040, 76426, 62880, 54107, 225044),
  size = c(320, 368, 100, 72, 82, 243),
  has_water= c(TRUE, TRUE, TRUE, TRUE, TRUE, TRUE)
)

colnames(cities_data) <- c('City', 'Population', 'size', 'has_water')

str(cities_data)
summary(cities_data)

cities_data$Density <- cities_data$Population / cities_data$size

cities_data

```
title: TASK 2 - Subsetting & Missing Data
author: "Lucia Bušová"
date: "9/11/2025"

### 2.1. Querying Vectors

```{r}
# code for solving task
# 2.1.1. Sequence vector
# 3th element
third_el <- s_vector[3]
third_el

# 2 to 5
two_five <- s_vector[2:5]
two_five

#greater than 20
greater_20 <- s_vector[s_vector > 20]
greater_20

# 2.1.2. Logical vector
# true values
true_vl <- l_vector[l_vector]
true_vl

# index
true_pst <- which(l_vector)
true_pst
```

### 2.2 Querying Your "Lumeria" List
```{r}
# code for solving task
# Using the `$` operator
T_list$T_descript

# matrix [[]]
T_list[['m']]

# second element
T_list[["s_vector"]][2]

```

### 2.3. Investigating Your Data Frame
```{r}
# code for solving task
# first 3 rows
head(cities_data, 3)

# 4th row, 1st column
cities_data[4, 1]

# contains logical value
cities_data$has_water
```

### 2.4. Handling "Corrupted" Data (`NA`)
```{r}
# code for solving task
#copying data
df_corrupted <- cities_data
#NA
df_corrupted[2, "size"] <- NA
df_corrupted[5, "Population"] <- NA
df_corrupted[6, "has_water"] <- NA
#count NA
sum(is.na(df_corrupted))
#count column NA
colSums(is.na(df_corrupted))
#mean of population with NA
mean(df_corrupted$Population)
#mean of population without NA
mean(df_corrupted$Population, na.rm = TRUE)
#remove NA
df_clean <- na.omit(df_corrupted)
df_clean
```

### 2.5. Bonus Challenge which
```{r}
# code for solving task
#highest population
highest_pop_index <- which.max(cities_data$Population)
cities_data[highest_pop_index, ]
#lowest density
lowest_density_index <- which.min(cities_data$Density)
cities_data[lowest_density_index, ]
```


